\section{Ultrasound}
Ultrasound (and all other types of sounds) are acoustic waves, i.e. a mechanical perturbation traveling in a medium. The acoustic frequency range:

\begin{itemize}
	\item Infrasound: < 20 Hz
	\item Audible sound: 20 Hz - 20 kHz
	\item Ultrasound: > 20 kHz 
	\item (Medical ultrasound: 1 - 20 MHz) 
\end{itemize}

Ultrasound is therefore mechanical oscillations with a frequency above 20 kHz that propagates through a medium that is elastic. Used for finding fungus in trees, used by bats to find their prey and in quality control of weld joint for finding cracks to give a few examples. The discovery of the piezoelectric property made this technique possible. 

	\subsection*{Ultrasound in medical imaging}
	Ultrasound is one of the most used imaging techniques in medicine and have an annual market growth of 3-4\%.Pros for this technique are that is non ionizing, need minimal safety requirements, its real time, can be made portable and is quite low cost. The downside with real time is the need of training personnel to understand/interpret the image while making the examination. Signal to noise ratio is quite low compared to other techniques and it need a acoustic window to operate (can't "see through" bone). 

	\subsection*{Probe - probing mechanism}
	Ultrasound is a longitudinal wave, variation of pressure that propagates, this mean that it needs a medium to travel through. The speed of sound is a function of the medium:

		\begin{equation}
			c = \sqrt{\frac{K} {\rho} }		
		\end{equation}

	Where K is the stiffness, stiffer the faster and $\rho$ is density: denser the slower. Like electrical circuits with resistance and current, the propagation of sound can be calculated with the Acoustic impedance:

		\begin{equation}
			Z = \rho c
		\end{equation}

\newpage

		\begin{table}[ht!]
		\centering
		\begin{tabular}{lllm{5cm}}\hline
		\textbf{Material} & bf\textbf{Density (kg$\text{m}^{-3}$)}& \textbf{Speed of sound}& \textbf{Acoustic impedance \newline (kg$\text{m}^{-2}$$\text{s}^{-1} \times 10^6$ )} \\

		Air & 1.3 &  330& 0.000429\\

		Water & 1000 & 1450& 1.50\\

		Bone & 1500 & 4000& 6.0\\
		
		Blood & 1060 & 1570& 1.59\\

		Muscle (average) & 1075 & 1590& 1.70 \\ \hline
		
		\end{tabular}
		\caption{Example of a table with line breaks}
		\label{tab:example}
		\end{table}

	The \textbf{wavelength} $\begin{bmatrix} m \end{bmatrix}$ is determined by: \textbf{Frequency} $\begin{bmatrix} Hz \end{bmatrix}$ (depends on the source) and \textbf{Sound-speed} $\begin{bmatrix} m/s \end{bmatrix}$ (depends on medium). 

	\subsection*{Matter interaction}
	After the sound wave left the source (probe) it interact with the material with the tissue, being \textbf{back-scattered}, \textbf{reflected}  and \textbf{refracted}. The sound wave is attenuated while moving through the tissue. 

	When the wave travels into a new medium the impedance of the medium and the incidence angle determines what happens to the wave (Huygens principle). If the wavefront is perpendicular to the medium the wavelength will change. If there is an oblique incidence angle it \textbf{refracts} giving a \textbf{reflection} against the new medium and \textbf{transmitted} wave through the new medium. The intensity of each part is dependent on: the \textbf{acoustic impedance difference} between the two materials and the \textbf{angle of incidence}.

	\textbf{Scattering} is the directionless reemission of incident energy caused by local inhomogeneities. Inside tissue, this scattering is caused by "tiny" particles, cells, large proteins, calcifications and so on. The interaction between this scattered waves is what causes the noisy speckle texture pattern in ultrasound images. 

	\textbf{Attenuation} is the loss of energy of the wave in the material. This energy is lost over distance due to \textbf{absorption}, \textbf{scattering}, mode conversions \textcolor{red}{?} and more. The amplitude decay increases with the depth (z) as a function of \textbf{some tissue properties} ($a_0$) and \textbf{acoustic frequency)}(f) 


		\begin{equation}
		A(z) = A_0 e^{-a_0 f^{n}z}
		\end{equation}

		\begin{wbox}{The tradeoff}
			\textcolor{OliveGreen}{The higher the frequency, the better the imaging resolution } \\
			\textcolor{red}{But, the worse the penetration depth} 
		\end{wbox}
	
	\subsection*{Piezo-electric effect}
	The purpose of the ultrasound transducer is to convert between electrical signal and acoustic energy. It act as both speaker and a microphone. By first emitting a very short sound pulse and then listen to the returning echoes (for a longer time then the emitting time). It is only possible to to one at a time. 

	The piezo-electric effect is a property that certain solids can have (some crystals and ceramics). Expose the material for electricity and there will respond with mechanical stress (like a speaker). Swapping places and expose it for mechanical stress will produce an electrical current (like a microphone)

	\subsection*{Image production}
	Producing the image involves switching between Transmit (Tx) mode : active (speaker) and Receive (Rx) mode : passive (microphone). There are different modes for the imaging also:

		\begin{itemize}
			\item \textbf{A(mplitude)-mode:} one scan line
			\item \textbf{B(rightness)-mode:} 2D spatial images with multiple neighboring A-lines
			\item \textbf{M(otion)-mode:} sequences of A-modes in time, thus 2D (popular in cardiology) 
		\end{itemize}

	Beamforming is the technique used to form sensor arrays for directional transmission and reception. By adjusting a time delay between element it is possible to mimic the spatially focus of a curved transducer. 

	\textcolor{red}{insert US focus illustration here} 


	\subsection*{Image quality}
	Contrast is what assesses visibility, so by comparing the image appearance between two distinct regions we want to be able to distinguish between object and background. The spatial resolution is the minimum distance at which is possible to distinguish two distinct point individually. The resolution of the ultrasound imaging system can be separated into \textbf{Axial resolution} and \textbf{Lateral resolution}. Axial resolution in which separating two reflectors or scatterers along the beams axis. This resolution depends on the length of the pulse, which need to be short for good axial resolution but needs to be long to deliver sufficient power for good signal to noise ratio. The lateral resolution separating reflectors side by side. This resolution depends on the focusing/beamforming. 

	\textbf{Artifacts} in ultrasound images are not uncommon and can even be used to interpret underlying causes some times. Examples of artifacts are: \textbf{Shadowing, speed-of-sound abberation, mirroring, reverberations and refraction} 

	\subsection*{Safety}
	In comparison with many other imaging systems in medicine, minimal. There are however four different metrics of safety that is regulated: 

		\begin{itemize}
			\item \textbf{Thermal index} (TI): quantifies/limit the acoustic power to avoid overheating the tissue internally
			\item \textbf{Surface temperature}: limit how hot the probe can be, to avoid burn injuries 
			\item \textbf{Spatial peak temporal average} (SPTA): limit the cumulative power at focus over time
			\item \textbf{Mechanical index} (MI):  limit peak negative acoustic pressure to avoid cavitation and streaming
		\end{itemize}

	\subsection*{Other ultrasound imaging techniques}
	With \textbf{Harmonic (b-mode) imaging} method we transmit a signal at a frequency f and then listen at frequency 2 times f. Comparing this to the fundamental B-mode we here get better resolution and reduced noise. \textbf{Doppler} technique uses the shift in of frequency relative to the motion of the object, compare to red and blue light shift in optics. With this technique its possible to quantitatively predict the blood flow velocity for an example. \textbf{Functional ultrasound} (fUS) can be seen as an extension of Doppler imaging and is used to examine brain activity (blood flow in the brain) with high frame-rate. \textbf{Elastography}, due to pathological process there may be changes in stiffness of the tissue, this technique estimate the tissue biomechanical characteristics. \textbf{Strain imaging}, since a stiff tissue compresses less it possible to detect differences inside a tissue by compressing while imaging, then measure local displacement and look at relative values/axial derivative \textcolor{red}{?}. \textbf{Acoustic Radiation Force} (ARF) uses focused acoustic energy to exert force in tissue by pushing it, remotely and locally, also known as "remote palpation". \textbf{Shear-wave Elastography} (SWE), here ultrasound is used to generate a \textbf{shear-wave} and then the temporal propagation is observed on consecutive ultrasound frames to infer the shear-wave speed. Stiffness difference in tissue can be observed that may indicate pathology. \textbf{Ultrasound computed tomography}. \textbf{Optoacoustic imaging} combines the use of the light pulses and ultrasound. Here is a light pulse transmitted and ultrasound received. The thermal expansion in the tissue creates the sound.  
















	


